Competing Constructions for German Measure Noun Phrases: from Usage Data to Experiment

### 2017-05-01 Latex/article.tex
Minor fixes after submission.

For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the findings with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
\footnote{In its colloquial meaning, the verb \textit{deduce} seems appropriate here.
However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
It would be highly surprising if this were possible and the same holds for experimental methods.
Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
Rather, under the standard approach, a theory of cognitive representation is pre-specified.
Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
Technical problems of statistical inference aside (see also Section~\ref{sec:rightstatistics}), a successful experiment corroborates the theory, and a failed experiment weakens the theory.
The simplest assumption would be that the same approach is adequate for corpus studies.

The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}, \citealp{Gries2003}).
It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
The similarity of a given feature vector (in a concrete sentence where one of the alternants is used) influences to a large extent which alternant is chosen by speakers.
This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
Prototype Theory and Exemplar Theory model essentially the same types off effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
\footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
Prototype Theory (with feature abstractions) is preferred here simply because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.

@@

The alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the PGC in (\ref{ex:intro:alternation1}) and the NAC in (\ref{ex:intro:alternation2}) are equally acceptable.
They are by-and-large functionally and semantically equivalent.
However, Section~\ref{sec:analyses} is devoted to developing hypotheses about subtle differences between them.%
\footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
See \cite{Hentschel1993,Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
As the usage and experimental data presented below (especially Section~\ref{sec:exp:fc}) should render it clear, there might be preferences, but we cannot assume either construction to be unacceptable.
}

@@

As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs.
They cannot be used in MNPs.}
%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
%A non-inflected version \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected version \textit{zwei Gläser roter Wein} co-exist.
%The non-inflected version denotes the quantity of wine fitting into two glasses, the inflected version denotes two glasses filled with wine.}
I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed.

@@

This section briefly reviews existing analyses of the \PGCa vs.\ \NACa\ alternation and related issues.
I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
I discuss four main aspects:
first, the different syntactic structures of the alternants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
third, a semantic prototypicality effect of the degrees of grammaticalisation (a preference to occur with cardinals);
fourth, preferences in different registers.

@@

Third, as described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
Both types of relations can be numerically quantified -- inasmuch as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
However, it is much more prototypical of quantities to be specified with numerical precision.
This is most obvious with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very prototypically used with exact numeral instead of unspecific quantifiers, although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
For illustration, (\ref{ex:determiners}) shows the expected alternants under this hypothesis.

@@

In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
Especially with abstract physical measure nouns (like \textit{centilitre}), exact numerical quantification is invited.
The statistical model presented in Section~\ref{sec:corpusstudies} will be specified in a way that it can detect such a preference.

@@

For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full,Schaefer2015b} and \citealp{BiemannEa2013,SchaeferBildhauer2013} for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
\footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
At the time of this writing, a newer 2016 version DECOW16 has already been released.}
I chose this corpus for two main reasons.%
\footnote{The use of web data for linguistic research does require explicit and careful justification.
Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
Clearly, using search engine results is ``bad science'' for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
\footnote{It was shown in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW.}
Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014,VanGoethemHuening2015,MuellerS2014,Schaefer2016c,SchaeferSayatz2014,SchaeferSayatz2016,Zimmer2015}. 
Therefore, DECOW is the obvious choice for this study.

@@

In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for nouns with a frequency up to 100 and a sample of 100 of all those with higher frequency.
The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
Given the careful sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
The sample is distributed in the data package accompanying this paper, however.
}

@@

The first and second point cannot be discussed in detail here.%
\footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
With regard to the third point, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
This does not do justice to frequentist methods in that mere rejection of the null hypothesis is characteristic only of Fisher's approach.
In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
Remarkably, given this statement, she does \textit{not} use informative priors and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
\footnote{In the words of \cite{Senn2011}: ``You may believe you are a Bayesian but you are probably wrong.''
\citet[347--348]{GelmanHill2006} ``view any noninformative prior distribution as inherently provisional'' and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
What is more, Bayesian estimators are in fact based on the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
With a modest model structure including 17 fixed effects and 2,646 data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
% Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
% However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.

@@

Finally, I show that MCMC does not necessarily lead to different results.
Actually, given the low complexity of the model and the large sample size, it would be surprising if they did (see Section~\ref{sec:rightstatistics}).
The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
The algorithm was run with 4 chains and 1,000 iterations and I used plausible default priors.
Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
\footnote{Consider that with a coefficient of 10, each increase by 1 in the regressor variable means an increase in odds of $exp(10)=22,026.47$.
To reliably estimate such coefficients, extremely large samples would be required.}
The algorithm converged and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
The resulting coefficients and intervals as well as an * for ``confidence interval does not contain 0'' are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size.
The signs and magnitudes of the coefficients are identical and confidence intervals have the same width and symmetry properties.
Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
Since both estimators converge, I only interpret the MLE model in the next section.

@@

The first experiment tests preferences for constructions explicitly.
\cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task \citep{Rosenbach2013}, where participants have to choose one of two alternants.
The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge, which is expected in the long run if a probability is turned into a binary decision.
I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants.
Participants had to choose between two sentences that differed only in that one contained the \NACa\ and the other contained the \PGCa.
The analysis compares the probabilities assigned to stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.

@@

The pairs of stimuli were the sentence containing the preferred construction (according to the corpus GLMM) and a modified version containing the dispreferred construction.
They were presented next to each other and a 20 second time limit for each choice was set.%
\footnote{No participant ever exceeded the time limit.}
The position on the screen (left\slash right) and the order of sentences were randomised for each participant.
As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
Thus, participants saw 39 pairs of sentences and 78 sentences in total.
They were instructed to select from each pair of sentences the one that seemed more natural to them in the sense that they would use it rather than the other one.
The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.

@@

\begin{acknowledgement}
  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, and Ulrike Sayatz for valuable discussions and comments.
  Elizabeth thankfully also fixed my English.
  Furthermore, I thank Ulrike for helping me to recruit the participants for the experiments.
  Finally, I am grateful to Kim Maser for her work on the annotation of the concordances.
  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
\end{acknowledgement}

>>>>

For such probabilistic generalisations estimated on corpus data, there has always been an interest in correlating the findings with results from experimental work (for example, \citealp{ArppeJaervikivi2007,BresnanEa2007,BresnanFord2010,DivjakGries2008,DivjakEa2016,FordBresnan2013}).
This often called a \textit{validation} of the corpus-derived findings, but \citet[303]{Divjak2016a} rightly criticises this choice of words ``because it creates the impression that behavioral experimental data is inherently more valuable than textual data,'' citing \cite{TummersEa2005}, who state that a corpus is ``a sample of spontaneous language use that is (generally) realized by native speakers.''
However, as \citet[486--487]{Dabrowska2016} convincingly argues, this does not imply that we can in some way ``deduce mental representations from patterns of use,'' \ie\ from corpus data.%
\footnote{In its colloquial meaning, the verb \textit{deduce} seems appropriate here.
However, in terms of the scientific method, we are most likely talking about \textit{induction} rather than \textit{deduction}.}
It would be highly surprising if this were possible and the same holds for experimental methods.
Nobody assumes that we can inductively infer mental representations from experiments, which -- as opposed to corpus studies -- even allow for direct access to the cognitive agent and offer much better possibilities to control experimental conditions and nuisance variables.
Rather, under the standard approach, a theory of cognitive representation is pre-specified.
Then, predictions are derived from this theory \textit{before} the experiment or the corpus study is conducted to \textit{test} the theory.
Technical problems of statistical inference aside, a successful experiment corroborates the theory, and a failed experiment weakens the theory.
The simplest assumption would be that the same approach is adequate for corpus studies.

The present study is conducted within the general paradigm of cognitive corpus linguistics and includes a comparison of the corpus findings with results from two experiments.
As a theoretical framework, I assume a model of similarity-based classification in the form of Prototype Theory.
Protoype Theory has been used in alternation research in cognitively oriented corpus linguistics (see \citealp{DivjakArppe2013}, \citealp{Gries2003}).
It is assumed that the variables annotated in a corpus study on an alternation phenomenon define prototypes for the alternants.
The similarity of a given feature vector (in a concrete sentence where one of the alternants is used) influences to a large extent which alternant is chosen by speakers.
This entails that a variant of Prototype Theory with features is assumed \citep{Rosch1978}, instead of the monolithic prototypes of earlier versions.
Prototype Theory is well suited for modeling constructional choices but it is just one of several similarity-based theories of classification, the most prominent other framework being Exemplar Theory (\citealp{MedinSchaffer1978,Hintzman1986}; see \citealp{StormsEa2000} for a comparison of the theories in experimental settings).
Prototype Theory and Exemplar Theory model essentially the same types of effects and differ mainly in whether they assume higher-level abstractions as part of the cognitive representation (Prototype Theory) or try to do without them (Exemplar Theory).
I am sceptical that corpus analysis alone could ever decide which theory is more suitable, given that substantial doubt has been voiced whether even experimental methods are ultimately able to do so \citep{Barsalou1990}.%
\footnote{In \cite{DivjakArppe2013}, it was shown using corpus data how both models form a converging picture.}
Prototype Theory (with feature abstractions) is preferred here simply because it fits the established alternation modeling paradigm, which relies on features being weighted in statistical models.
In Section~\ref{sec:analyses}, a prototype-theoretical parlance is therefore adopted.

@@

The alternation can be observed \textit{only when the kind noun occurs with an attributive adjective but without a determiner}, as in (\ref{ex:intro:alternation}), where both the PGC in (\ref{ex:intro:alternation1}) and the NAC in (\ref{ex:intro:alternation2}) are equally acceptable.
They are by-and-large functionally and semantically equivalent.
However, Section~\ref{sec:analyses} is devoted to developing hypotheses about subtle differences between them.%
\footnote{Some descriptive and normative grammars take stronger positions with regard to the acceptability of the two options.
See \cite{Hentschel1993} and \cite{Zimmer2015} for analyses of the sometimes absurd stances taken in grammars of German.
As the usage and experimental data presented below (especially Section~\ref{sec:exp:fc}) should render it clear, there might be preferences, but we cannot assume either construction to be unacceptable.
}

@@

As for (\ref{it:intro:noifl}), some measure nouns have been grammaticalised in a way that they always appear in their non-inflected form.
They are typical measure nouns like \textit{Gramm} `gram', \textit{Pfund} `pound' or \textit{Prozent} `percent', which have no normal plural forms at all.%
\footnote{Plurals like \textit{Pfunde} and \textit{Prozente} have special meanings and have very restricted uses, mostly in idiomatic expressions and prefabs such as \textit{Pfunde verlieren} `lose pounds.'
They cannot be used in normal MNPs.}
%\footnote{With neglectable frequency, this effect is extended to container nouns like \textit{Glas} `glass', but with a twist.
%A non-inflected version \textit{zwei Glas roter Wein} `two glasses of wine' and an inflected version \textit{zwei Gläser roter Wein} co-exist.
%The non-inflected version denotes the quantity of wine fitting into two glasses, the inflected version denotes two glasses filled with wine.}
I treat these cases like other measure nouns because they enter into both the \NACa\ as in (\ref{ex:uninflectedmeasures:a}) and the \PGCa\ as in (\ref{ex:uninflectedmeasures:b}).
In Section~\ref{sec:analyses}, however, degrees of grammaticalisation as a factor influencing the alternation will be discussed prominently.

@@

This section briefly reviews existing analyses of the \PGCa\ vs.\ \NACa\ alternation and related issues.
I also develop my own analysis and the appropriate hypotheses for the empirical studies presented in Sections~\ref{sec:corpusstudies} and \ref{sec:experimental}.
I discuss four main aspects:
first, the different syntactic structures of the alternants \NACa\ and \PGCa\ in relation to their syntactic prototypes \NACb\ and \PGCd;
second, degrees of grammaticalisation of the measure noun associated with the two prototypes;
third, a semantic prototypicality effect of the degrees of grammaticalisation (a preference to occur with cardinals);
fourth, preferences in different registers.

@@

Third, as described above, the grammaticalisation path leads from NPs denoting individuated objects standing in a \textit{part-of} relation to a construction with a more diffuse \textit{quantity-of} relation.
Both types of relations can be numerically quantified -- inasmuch as a precise number of \textit{parts} or a numerically exact \textit{quantity} can be specified.
However, it is much more prototypical of quantities to be specified with numerical precision.
This is most obviously so with the highly grammaticalised physical measure nouns like \textit{centilitre} which are very prototypically used with exact numeral instead of unspecific quantifiers, although both options are available (\textit{three centilitres} vs.\ \textit{several centilitres}).
Since the \NACa\ is more closely associated with the \textit{quantity-of} relation, cardinals as attributes of the measure noun are expected to have a higher proportional frequency in the \NACa.
For illustration, (\ref{ex:determiners}) shows the expected alternants under this hypothesis.

@@

In (\ref{ex:determiners:nac}), the measure noun is modified by a cardinal \textit{drei} `three', and hence the \NACa\ is preferred.
In (\ref{ex:determiners:pgc}), the measure noun is modified by a non-cardinal determiner \textit{einige} `some', and the \PGCa\ is preferred.
Especially with abstract physical measure nouns (like \textit{centilitre}), exact numerical quantification is invited.
By hpothesis, however, this goes beyond a selection effect tied to measure lemmas and cardinal quantifiers are expected to co-occur relatively more often with the \NACa.
The statistical model presented in Section~\ref{sec:corpusstudies} will be specified in a way that it can detect such a preference.

@@

For the present study, I used the German \textit{Corpus from the Web} (COW) in its 2014 version DECOW14A (\citealp{SchaeferBildhauer2012full}, and \citealp{Schaefer2015b}, as well as \citealp{BiemannEa2013}, and \citealp{SchaeferBildhauer2013}, for overviews of web corpora in general and the methodology of their construction), which contains almost 21 billion tokens.%
\footnote{The COW corpora (Dutch, English, French, German, Spanish, Swedish) are made available for free at \url{https://www.webcorpora.org}.
At the time of this writing, a newer 2016 version DECOW16 has already been released.}
I chose this corpus for two main reasons.%
\footnote{The use of web data for linguistic research does require explicit and careful justification.
Due to the noisy nature and unknown composition of the web, only carefully designed and established web corpora like the COW corpora or the SketchEngine corpora \citep{KilgarriffEa2014} should be used.
Clearly, using search engine results is ``bad science'' for many reasons, most prominently total irreproducibility of results, as \cite{Kilgarriff2006} pointed out more than ten years ago.
Careless use of search engine results is still found, however, see for example \citet[171--175]{DeclerckBrems2016}.}
First, the external validity of any study is increased through a higher heterogeneity of the sample \citep[30]{MaxwellDelaney2004}, and the DECOW corpus has clearly a much more heterogeneous composition compared to the only other very large corpus of German, the DeReKo \citep{KupietzEa2010} of the Institute for the German Language (IDS), which contains almost exclusively newspaper texts.%
\footnote{It was shown in \cite{W16-2601} that, for example, the spread of topics is much smaller in DeReKo compared to DECOW.}
Second, it was already mentioned that normative grammars often adopt clear positions regarding the grammaticality of either the \NACa\ or the \PGCa.
Thus, newspaper text or any other text that conforms strongly to normative grammars might not represent the alternation phenomenon fully (and without bias) because authors and proofreaders who must adhere to normative guidelines might favour one alternative or the other explicitly.
Web corpora, on the other hand, contain at least some amount of non-standard language from forums and similar sources.
For these or similar reasons, COW corpora have been used in a number of peer-reviewed publications, for example \cite{VanGoethemHiligsmann2014}, \cite{VanGoethemHuening2015}, \cite{MuellerS2014}, \cite{Schaefer2016c}, \cite{SchaeferSayatz2014}, \cite{SchaeferSayatz2016}, and \cite{Zimmer2015}. 
Therefore, DECOW is the obvious choice for this study.

@@

In step (\ref{enum:data:step3}), each of these 2,365 noun–noun combinations was queried in the target constructions (\PGCa\ and \NACa) individually in each of the first ten slices of DECOW (roughly 10 billion tokens).
In order to reduce the sample size for the manual annotation process, the final concordance was sampled from the results of these 2,365 queries.
Since the mass nouns in the sample were distributed according to the usual power law, I used all hits for nouns with a frequency up to 100 and a sample of 100 of all those with higher frequency.
The final sample contained 6,843 sentences, which was reduced to 5,063 in the manual annotation process due to removal of noisy material, erroneous hits and uninformative cases where the measure noun was in the genitive, in which case the \NACa\ cannot be distinguished from the \PGCa.
Given the careful sampling procedure described in this section, we can be highly sure that it contains all relevant and reasonably frequent noun–noun combinations in the target constructions.%
\footnote{In a similar fashion, the 100 most frequent measure nouns occurring with plural kind nouns were listed and queried, resulting in a sample of 871 sentences.
As stated in Section~\ref{sec:germanmeasurenps}, the \NACa\ is virtually never used with plural kind nouns, and this sample was not used except for quantifying the frequency of occurrence of the constructions (67 times \NACa\ and 794 times \PGCa).
However, the sample is distributed in the data package accompanying this paper.
}

@@

The first and second point cannot be discussed in detail here.%
\footnote{However, there is a number of critical papers by statisticians (even prominent Bayesians) and philosophers of science in which the extreme credibility that has recently been assigned to (subjective and automatic) Bayesian approaches (and especially the standard inductivist interpretation of Bayesianism) is questioned (for example, \citealp{GelmanShalizi2013,Mayo2011,Senn2011}).}
With regard to the third point, \citet[251--252]{Levshina2016} argues for Bayesian estimation in mixed regression settings.
First, she claims that ``while frequentist statistics only allows one to test whether the null hypothesis can be rejected, Bayesian statistics enables one both to test the null hypothesis and to estimate the probability of specific parameter values given the data.''
This does not do justice to frequentist methods in that a strong focus on the rejection of the null hypothesis is characteristic only of Fisher's approach.
In the Neyman-Pearson approach, results ideally \textit{favour} the main hypothesis vis-à-vis the alternative hypothesis (cf.\ \citealp{Lehmann1993,Lehmann2011,Perezgonzalez2015}).
Also, especially Neyman-style frequentism has well-known extensions to estimation, for example in the form of confidence intervals (see \citealp{GreenlandEa2016}, esp.\ p.\ 340).
She then explains that a ``distinctive feature of Bayesian statistics is the use of so-called priors'' and that ``posterior probabilities depend on both the prior beliefs and the data, whereas the results of a frequentist model depend only on the data'' \cite[252]{Levshina2016}.
Remarkably, given this statement, she does \textit{not} use informative priors and in her footnote 8 \cite[252]{Levshina2016} admits that priors were probed using trial and error.
So, the proclaimed major advantage of Bayesian modeling was apparently not taken advantage of.%
\footnote{In the words of \cite{Senn2011}: ``You may believe you are a Bayesian but you are probably wrong.''
\citet[347--348]{GelmanHill2006} ``view any noninformative prior distribution as inherently provisional'' and give recommendations how to proceed once posteriors have been obtained from noninformative priors.}
Now, Maximum Likelihood Estimation (MLE) -- the traditional method which could have been used instead -- is not exactly \textit{frequentist} in the sense of Neyman-Pearson testing theory.
MLE, like inductive Bayesianism, conditions on the particular data inasmuch as it searches for the most likely set of parameters given the data.
What is more, Bayesian estimators are in fact based on the Likelihood and merely multiply it by the prior \citep[6--8]{GelmanEa2014}.
If the prior is flat, results converge (see also \citealp[347]{GelmanHill2006}).
The same is true if the sample size is large compared to the number of parameters, at least for finite-dimensional parameter models \citep[1119--1120]{Freedman1999}, a well-established result known as the \textit{Bernstein-von Mises theorem}.
With a modest model structure including 17 fixed effects and 2,646 data points in \cite{Levshina2016}, it is highly likely that the same results would have been obtained with Maximum Likelihood methods.
In fact, she admits that changing the priors did not lead to substantially different results in her footnote 8.
This is a clear sign that the prior is ``swamped by the data'' \citep[1119]{Freedman1999}.
% Going back to the argument that Bayesian methods are allegedly more cognitively plausible than `frequentist' methods, the claim might still be true.
% However, most similar studies (dating back to \citealp{BresnanEa2007}) use an estimator (MLE) that is not frequentist in a narrow sense, and which is as cognitively plausible as Bayesian estimators because it most likely produces the same results in these studies.

@@

Finally, I show that MCMC does not necessarily lead to different results.
Actually, given the low complexity of the model and the large sample size, it would be surprising if they did (see Section~\ref{sec:rightstatistics}).
The model was re-estimated using the \textit{stan\_glmer} function from the \textit{rstanarm} package, which provides an \textit{lme4}-compatible syntax for estimating common model types with the \textit{stan} software \citep{CarpenterEa2017}.
The algorithm was run with 4 chains and 1,000 iterations and I used plausible default priors.
Most notably, priors for coefficients were specified as $\mathcal{N}(0,10)$ because coefficients higher than 10 or lower than $-10$ are extremely rare in well-specified models on appropriate data.%
\footnote{Consider that with a coefficient of 10, each increase by 1 in the regressor variable increases the odds by $exp(10)=22,026.47$.
To reliably estimate such coefficients, extremely large samples would be required.}
The algorithm converged and for all coefficients, the $\hat{R}$ diagnostic was exactly 1.
The resulting coefficients and intervals as well as an * for ``confidence interval does not contain 0'' are also given in Table~\ref{tab:bigtable} in the columns labelled \textit{MCMC}.
Both methods lead to exactly the same results (minus negligible numerical differences) as expected given the modest complexity of the model structure and the large sample size.
The signs and magnitudes of the coefficients are identical and confidence intervals have the same width and symmetry properties.
Figure~\ref{fig:fixeffs} illustrates this by also showing both estimates.
Since both estimators converge, I only interpret the MLE model in the next section.

@@

The first experiment tests preferences for constructions explicitly.
\cite{FordBresnan2013} use the \textit{split-100} task in which participants have to distribute 100 points between the alternatives, assigning more points to more natural sounding alternative.
In essence, participants distribute a probability mass between two alternants, which is intended to produce more subtle results compared to a two-alternative forced-choice task such as in \cite{Rosenbach2013}, where participants have to choose one of two alternants.
The split-100 paradigm has been criticised in \cite{ArppeJaervikivi2007}.
The criticism was reiterated in \cite{DivjakEa2016}, where they use a forced-choice task.
In \cite{VerhoevenTemme2017}, it was shown that results from forced-choice and split-100 experiments mostly converge, which is expected in the long run if a probability is turned into a binary decision.
I present a forced-choice experiment in this Section and not a split-100 experiment, mainly because in a dry run of the experiment, participants complained about the unnaturalness of distributing a probability mass across two alternants.
Participants had to choose between two sentences that differed only in that one contained the \NACa\ and the other contained the \PGCa.
The analysis compares the probabilities assigned to stimuli by the corpus-derived model with the frequency with which participants chose the alternants for the same stimuli.

@@

The final pairs of stimuli were the sentence containing the attested and preferred alternant (according to the corpus GLMM) on the one hand and a modified version containing the dispreferred alternant on the other hand.
They were presented next to each other and a 20 second time limit for each choice was set.%
\footnote{No participant ever exceeded the time limit.}
The position on the screen (left\slash right) and the order of sentences were randomised for each participant.
As fillers, 23 pairs of sentences exemplifying similar but unrelated alternation phenomena from German morpho-syntax were used.
Thus, participants saw 39 pairs of sentences and 78 sentences in total.
They were instructed to select from each pair of sentences the one that seemed more natural to them in the sense that they would use it rather than the other one.
The experiment was conducted using \textit{PsychoPy} \citep{Peirce2007}.

@@

\begin{acknowledgement}
  I thank (in alphabetical order) Felix Bildhauer, Susanne Flach, Elizabeth Pankratz, Samuel Reichert, Ulrike Sayatz, and Christian Zimmer for valuable discussions and comments.
  Elizabeth thankfully also fixed my English.
  Furthermore, I thank Ulrike for helping me to recruit the participants for the experiments.
  Finally, I am grateful to Kim Maser for her work on the annotation of the concordances.
  The research presented here was made possible in part through funding from the \textit{Deutsche Forschungsgemeinschaft} (DFG, personal grant SCHA1916/1-1).
\end{acknowledgement}
